\selectlanguage{english}

\normalsize
\chapter{Acknowledgements}

This thesis was completed under Dr. Marija Mitrović Dankulov supervision at the Scientific Computing Laboratory at the Institute of Physics Belgrade. I want to express my sincere gratitude to my supervisor for her invaluable guidance, support, and patience during my studies. Her mentorship has been instrumental in helping me to complete my dissertation.  

I am grateful to the head of SCL, Dr. Antun Balaz, for his ongoing assistance and advice through all these years. I also want to thank colleagues from the laboratory and institute for making the workplace so enjoyable. I wish to acknowledge collaborators Dr. Aleksandra Alorić, Dr. Jelena Smiljanić, and Dr. Aleksandar Tomasević for their contributions to the research presented in this thesis. Their expertiese, valuable insights, and numerous discussions we had, helped me to refine my research. It has been my pleasure to collaborate with Darja Cvetković and have the opportunity to learn so much from her.

I thank my family and friends for their love and support. To my parents, who gave me tremendous encouragement and understanding, especially to my mom, for being by my side and believing in me.

The research presented here was supported by the Ministry of Education, Science, and Technological Development of the Republic of Serbia, the National Project ON171017 Modeling and Numerical Simulations of Complex Many-Body Systems; by the Science Fund of the Republic of Serbia, the Artificial intelligence theoretical foundations for advanced spatio-temporal modeling of data and processes (ATLAS) project; by Innovation Fund of Republic Serbia the Platform for REmote development of Autonomous Driving algorithms in a realistic environment (READ) project and by 60seconds startup. Numerical simulations were run on the PARADOX supercomputing facility at the Scientific Computing Laboratory of the Institute of Physics Belgrade.

\hfill

\justify
